---
title: "Using database"
author: ""
date: "02/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
dbListTables(db)

dbGetQuery(db, "SELECT * FROM demo LIMIT 5")
dbGetQuery(db, "select * from trump limit 5")
dbGetQuery(db, "select * from vac limit 5")

```


### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2021, and select county with less than 10 pct vaccination

```{r}
dbGetQuery(db, "SELECT *, MAX(Date) FROM vac GROUP BY Recip_County")

dbGetQuery(db, "SELECT Recip_County  FROM vac WHERE Date = '2021-08-01' AND Completeness_pct < 0.1") # matter of arguments (SELECT - FROM - WHERE) matters!

```
#### Ordering and aggregation

1. Order counties with the highest rate of vaccination on Aug 14, 2021
2. Calculate the state level vaccine count


```{r}
dbGetQuery(db, "SELECT Recip_County, Date, Completeness_pct FROM vac WHERE Date = '2021-08-14' ORDER BY Completeness_pct DESC")

dbGetQuery(db, "SELECT Recip_State, SUM(Series_Complete_Yes) FROM vac GROUP BY Recip_State")
```


### Merging tables

- merge vaccine and demo table by fips
- (advanced) calculate per 1000 cases and deaths

```{r}
dbGetQuery(db, "SELECT vac.* FROM vac LEFT JOIN demo ON vac.FIPS = demo.FIPS")

dbGetQuery(db, "SELECT *, Series_Complete_Yes/Census2019 FROM vac") # necessary vars are missing, but this is how calculations work

```

## dblyr

Now we will use `dbplyr`

### Get the object for each table

```{r}
library(dbplyr)
data_vac <- tbl(db, from = "vac")
data_demo <- tbl(db, from = "demo")
data_trump <- tbl(db, from = "trump")
```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
data_vac %>% head()
data_demo %>% head()
data_pres %>% head()
```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 14 Aug, 2021, and select county with more than 10 percent vaccination rate

```{r}
data_vac %>% slice_max(Date)
data_vac %>% filter(Date == "2021-08-14", Completeness_pct > 10) %>% select(Recip_County)
```
#### Ordering and aggregation

1. Order counties with the highest number of vaccine percentage
2. Calculate the state level vaccination total


```{r}
data_vac %>% arrange(desc(Completeness_pct))

data_vac %>% group_by(Recip_State) %>% summarise(vaccination = mean(Completeness_pct))

```


### Merging tables

- merge vaccine and demo table by fips
- calculate vaccination rate manually from Series_Complete_Yes and TotalPop
  - compare with Series_Complete_Pop_Pct


```{r}
data_vac %>% inner_join(data_demo, by = "FIPS")

data_vac %>% select(Series_Complete_Yes, Census2019, Series_Complete_Pop_Pct) %>% mutate(vacc_rate = Series_Complete_Yes/Census2019)

```
